---
title: "f1000 R codes for upload"
author: "Amirsalar Jafari Gorizi"
date: "2026-01-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Set directory
```{r}
out_dir <- "/Users/Downloads/"
```

## Libraries
```{r}
# ---- Packages: install if missing, then load ----
req <- c(
  "dplyr",
  "stringr",
  "strucchange",
  "ggplot2",
  "tidyr",
  "readr",
  "lubridate",
  "CausalImpact",
  "zoo",
  "forecast",
  "tseries"
)

to_install <- setdiff(req, rownames(installed.packages()))
if (length(to_install) > 0) {
  install.packages(to_install, repos = "https://cloud.r-project.org")
}

invisible(lapply(req, library, character.only = TRUE))


```

# Bai-Parron stylistic constructs combined chart
```{r cars}
# 1. Load your data
scores <- read.csv("scores.csv")

# 2. Define your variables and formatted labels
vars <- c("share_flag_fdg", "share_flag_lrr", "share_flag_imbd")

# 3. Initialize data frames to store results for plotting
plot_data_raw    <- data.frame() # NEW: To store the raw time series
plot_data_fitted <- data.frame() # To store the step-function lines
plot_breaks      <- data.frame() # To store the vertical break lines

# 4. Loop through each variable to calculate breaks
for (v in vars) {
  if (!v %in% names(scores)) {
    cat("⚠️ Variable not found:", v, "\n")
    next
  }
  
  # Prepare Time Series
  data_series <- scores[[v]]
  data_ts <- ts(data_series, start = 2012, frequency = 1)
  
  # --- Step A: Save Raw Data for Plotting ---
  temp_raw <- data.frame(
    year = as.numeric(time(data_ts)),
    value = as.numeric(data_series),
    variable = v
  )
  plot_data_raw <- rbind(plot_data_raw, temp_raw)
  
  # --- Step B: Fit Bai–Perron model ---
  bp_model <- try(breakpoints(data_ts ~ 1, h = 0.15), silent = TRUE)
  
  if (inherits(bp_model, "try-error")) {
    cat("❌ Model failed for:", v, "\n")
    next
  }
  
  # --- Step C: Determine optimal number of breaks (BIC) ---
  if (!is.null(bp_model$BIC) && length(bp_model$BIC) > 0) {
    opt_breaks <- which.min(bp_model$BIC) - 1
  } else {
    opt_breaks <- 0
  }
  
  # --- Step D: Refit final model ---
  if (opt_breaks > 0) {
    bp_final <- breakpoints(data_ts ~ 1, breaks = opt_breaks)
  } else {
    bp_final <- bp_model
  }
  
  # --- Step E: Extract Fitted Data ---
  fitted_vals <- fitted(bp_final)
  temp_fit <- data.frame(
    year = as.numeric(time(data_ts)),
    value = as.numeric(fitted_vals),
    variable = v
  )
  plot_data_fitted <- rbind(plot_data_fitted, temp_fit)
  
  # --- Step F: Extract Breakpoints (if any) ---
  if (!is.null(bp_final$breakpoints) && any(!is.na(bp_final$breakpoints))) {
    break_years <- time(data_ts)[bp_final$breakpoints]
    temp_breaks <- data.frame(
      break_year = as.numeric(break_years),
      variable = v
    )
    plot_breaks <- rbind(plot_breaks, temp_breaks)
  }
}

# 5. Create the Combined Plot
p <- ggplot() +
  
  # LAYER 1: Raw Data (Faint lines)
  geom_line(data = plot_data_raw, 
            aes(x = year, y = value, color = variable), 
            alpha = 0.4, linewidth = 0.8, linetype = "solid") +
  
  # LAYER 2: Fitted Trends (Bold Step Lines)
  geom_line(data = plot_data_fitted, 
            aes(x = year, y = value, color = variable), 
            linewidth = 1.2) +
  
  geom_vline(data = plot_breaks %>% 
               mutate(nudge = case_when(
                 variable == "share_flag_fdg"  ~ -0.12,
                 variable == "share_flag_lrr"  ~ 0,
                 variable == "share_flag_imbd" ~ 0.12,
                 TRUE ~ 0
               )), 
             aes(xintercept = break_year + nudge, color = variable), 
             linetype = "dashed", linewidth = 0.8, alpha = 0.8) +
  
  # Styling and Legend
  scale_x_continuous(breaks = 2012:2025) + 
  scale_color_manual(
    # 1. Define the specific variables
    breaks = c("share_flag_fdg", "share_flag_lrr", "share_flag_imbd"),
    
    # 2. Assign the exact colors you requested
    values = c("share_flag_fdg" = "#4c72b0",  # Blue (Homogeneity)
               "share_flag_lrr" = "#dd8452",  # Orange (Conventionality)
               "share_flag_imbd" = "#55a868"), # Green (Predictability)
               
    # 3. Label them in the legend
    labels = c("Syntactic Homogeneity", "Lexical Conventionality", "Rhythmic Predictability"),
    name = "Indicator"
  ) +
  labs(
    title = "Structural Breaks in Flagged Proportion of Abstracts",
    subtitle = "Faint lines = Raw scores; Bold lines = Structural means, Vertical dashed lines = Structural breaks",
    y = "Share Value",
    x = "Year"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    panel.grid.minor = element_blank(),
    plot.title = element_text(face="bold")
  )

# 6. Output
print(p)

ggsave("bai_perron_scores.png", plot = p, width = 10, height = 6)
```

## Bai-Parron stylistic constructs combined "Infectious Disease & Crisis" & "General Scientific Corpus"
```{r}

# 1. Load your raw data
# Assuming your raw dataset is named 'raw_data'
raw_data <- read.csv("/Users/Downloads/f1000_unique_scored.csv")

# 2. Categorize
crisis_keywords <- "covid|sars-cov-2|coronavirus|pandemic|epidemic|outbreak|h1n1|zika|ebola|influenza|virus|pathogen|infectious disease|public health emergency"

processed_data <- raw_data %>%
  mutate(category = ifelse(
    str_detect(tolower(title), crisis_keywords) | 
    str_detect(tolower(abstract), crisis_keywords),
    "Infectious Disease & Crisis",  # The Control Group (Historically Rigid?)
    "General Scientific Corpus"     # The Treatment Group (GPT-3 Affected?)
  ))

# 3. Calculate Yearly Shares (Proportions) for the flag columns
scores_long <- processed_data %>%
  group_by(publication_year, category) %>%
  summarise(
    share_flag_fdg = mean(flag_llm_fdg, na.rm = TRUE),
    share_flag_lrr = mean(flag_llm_lrr, na.rm = TRUE),
    share_flag_imbd = mean(flag_llm_imbd, na.rm = TRUE),
    .groups = 'drop'
  )

# 4. Initialize storage for Bai-Perron results
vars <- c("share_flag_fdg", "share_flag_lrr", "share_flag_imbd")
plot_data_raw    <- data.frame()
plot_data_fitted <- data.frame()
plot_breaks      <- data.frame()

# 5. Loop through Categories and Variables to calculate breaks
categories <- unique(scores_long$category)

for (cat_name in categories) {
  cat_data <- scores_long %>% filter(category == cat_name)
  
  for (v in vars) {
    # Prepare Time Series for this category
    data_series <- cat_data[[v]]
    # Adjust start year based on your data range for that specific category
    start_yr <- min(cat_data$publication_year)
    data_ts <- ts(data_series, start = start_yr, frequency = 1)
    
    # Save Raw Data
    temp_raw <- data.frame(
      year = as.numeric(time(data_ts)),
      value = as.numeric(data_series),
      variable = v,
      category = cat_name
    )
    plot_data_raw <- rbind(plot_data_raw, temp_raw)
    
    # Fit Bai–Perron model
    # h=0.15 ensures we have enough observations between breaks
    bp_model <- try(breakpoints(data_ts ~ 1, h = 0.15), silent = TRUE)
    
    if (inherits(bp_model, "try-error")) next
    
    # Determine optimal number of breaks (BIC)
    opt_breaks <- if(!is.null(bp_model$BIC)) which.min(bp_model$BIC) - 1 else 0
    
    # Refit final model
    bp_final <- if(opt_breaks > 0) breakpoints(data_ts ~ 1, breaks = opt_breaks) else bp_model
    
    # Extract Fitted Data
    fitted_vals <- fitted(bp_final)
    temp_fit <- data.frame(
      year = as.numeric(time(data_ts)),
      value = as.numeric(fitted_vals),
      variable = v,
      category = cat_name
    )
    plot_data_fitted <- rbind(plot_data_fitted, temp_fit)
    
    # Extract Breakpoints
    if (!is.null(bp_final$breakpoints) && any(!is.na(bp_final$breakpoints))) {
      break_years <- time(data_ts)[bp_final$breakpoints]
      temp_breaks <- data.frame(
        break_year = as.numeric(break_years),
        variable = v,
        category = cat_name
      )
      plot_breaks <- rbind(plot_breaks, temp_breaks)
    }
  }
}

# 6. Create the Faceted Plot
p <- ggplot() +
  # LAYER 1: Raw Data
  geom_line(data = plot_data_raw, 
            aes(x = year, y = value, color = variable), 
            alpha = 0.3, linewidth = 0.7) +
  
  # LAYER 2: Fitted Trends
  geom_line(data = plot_data_fitted, 
            aes(x = year, y = value, color = variable), 
            linewidth = 1.2) +
  
  # LAYER 3: Vertical Break Lines (with category-specific nudging)
  geom_vline(data = plot_breaks %>% 
               mutate(nudge = case_when(
                 variable == "share_flag_fdg"  ~ -0.15,
                 variable == "share_flag_lrr"  ~ 0,
                 variable == "share_flag_imbd" ~ 0.15,
                 TRUE ~ 0
               )), 
             aes(xintercept = break_year + nudge, color = variable), 
             linetype = "dashed", linewidth = 0.6, alpha = 0.7) +
  
  # Faceting by Category
  facet_wrap(~category, scales = "free_y") +
  
  # Styling
  scale_x_continuous(breaks = seq(2012, 2025, by = 2)) + 
  scale_color_manual(
    values = c("share_flag_fdg" = "#4c72b0", 
               "share_flag_lrr" = "#dd8452", 
               "share_flag_imbd" = "#55a868"),
    labels = c("Fast-DetectGPT (FDG)", "Log-Rank (LRR)", "IMBD Classifier"),
    name = "LLM Detector"
  ) +
  labs(
    y = "Proportion of Abstracts Flagged",
    x = "Year"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom", plot.title = element_text(face="bold"))

print(p)
```

## Bai-Parron reviewer recommendations combined chart
```{r}
# 1. Load data
recommendations <- read.csv(
  "/Users/Downloads/f1000_unique_scored.csv"
)

# 2. Compute yearly percentages (overall, no categories)
recommendations_trend <- recommendations %>%
  group_by(publication_year, reviewer_recommendation) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(publication_year) %>%
  mutate(percentage = (count / sum(count)) * 100) %>%
  ungroup()

# 3. Setup
vars <- c("reject", "approve", "approve-with-reservations")

plot_data_raw    <- data.frame()
plot_data_fitted <- data.frame()
plot_breaks      <- data.frame()

# 4. Loop through recommendation types
for (v in vars) {

  v_data <- recommendations_trend %>%
    filter(reviewer_recommendation == v) %>%
    arrange(publication_year)

  # Skip if too little data
  if (nrow(v_data) < 5) next

  # Create time series
  data_ts <- ts(
    v_data$percentage,
    start = min(v_data$publication_year),
    frequency = 1
  )

  # --- Save raw data ---
  plot_data_raw <- rbind(
    plot_data_raw,
    data.frame(
      year = as.numeric(time(data_ts)),
      value = as.numeric(data_ts),
      variable = v
    )
  )

  # --- Bai–Perron model ---
  bp_model <- try(breakpoints(data_ts ~ 1, h = 0.15), silent = TRUE)
  if (inherits(bp_model, "try-error")) next

  opt_breaks <- if (!is.null(bp_model$BIC)) {
    which.min(bp_model$BIC) - 1
  } else {
    0
  }

  bp_final <- if (opt_breaks > 0) {
    breakpoints(data_ts ~ 1, breaks = opt_breaks)
  } else {
    bp_model
  }

  # --- Fitted values ---
  plot_data_fitted <- rbind(
    plot_data_fitted,
    data.frame(
      year = as.numeric(time(data_ts)),
      value = as.numeric(fitted(bp_final)),
      variable = v
    )
  )

  # --- Breakpoints ---
  if (!is.null(bp_final$breakpoints) && any(!is.na(bp_final$breakpoints))) {
    plot_breaks <- rbind(
      plot_breaks,
      data.frame(
        break_year = as.numeric(time(data_ts)[bp_final$breakpoints]),
        variable = v
      )
    )
  }
}

# 5. Plot
p <- ggplot() +
  geom_line(
    data = plot_data_raw,
    aes(x = year, y = value, color = variable),
    alpha = 0.35,
    linewidth = 0.8
  ) +
  geom_line(
    data = plot_data_fitted,
    aes(x = year, y = value, color = variable),
    linewidth = 1.2
  ) +
  geom_vline(
    data = plot_breaks,
    aes(xintercept = break_year, color = variable),
    linetype = "dashed",
    alpha = 0.8
  ) +
  scale_color_manual(
    values = c(
      "reject" = "#e74c3c",
      "approve" = "#2ecc71",
      "approve-with-reservations" = "#3498db"
    ),
    name = "Recommendation"
  ) +
  labs(
    title = "Structural Breaks in Reviewer Recommendations",
    subtitle = "Raw yearly percentages and Bai–Perron step means",
    x = "Publication Year",
    y = "Percentage (%)"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

print(p)


# Optional: Save
ggsave("/Users//Documents/bai_perron_reviews.png", plot = p, width = 10, height = 6)
```

## Bai-Parron reviewer recommendations combined "Infectious Disease & Crisis" & "General Scientific Corpus"
```{r}
# 1. Load your raw data
# Ensure your CSV has: publication_year, reviewer_recommendation, and abstract
raw_data <- read.csv("/Users/Downloads/f1000_unique_scored.csv")

# 2. Categorize based on Abstract Keywords
crisis_keywords <- "covid|sars-cov-2|coronavirus|pandemic|epidemic|outbreak|h1n1|zika|ebola|influenza|virus|pathogen|infectious disease|public health emergency"

processed_data <- raw_data %>%
  mutate(category = ifelse(
    str_detect(tolower(abstract), crisis_keywords),
    "Infectious Disease & Crisis", 
    "General Scientific Corpus"
  ))

# 3. Calculate Yearly Percentages for Recommendations per Category
# This matches your 'trend_percent' logic: (count / yearly total) * 100
recommendations_trend <- processed_data %>%
  group_by(publication_year, category, reviewer_recommendation) %>%
  summarise(count = n(), .groups = 'drop') %>%
  group_by(publication_year, category) %>%
  mutate(percentage = (count / sum(count)) * 100) %>%
  ungroup()

# 4. Initialize storage for plotting
vars <- c("approve", "reject", "approve-with-reservations")
plot_data_raw    <- data.frame()
plot_data_fitted <- data.frame()
plot_breaks      <- data.frame()
categories       <- unique(recommendations_trend$category)

# 5. Loop through Categories and Recommendation Types
for (cat_name in categories) {
  for (v in vars) {
    # Filter data for specific recommendation type in specific group
    cat_v_data <- recommendations_trend %>% 
      filter(category == cat_name, reviewer_recommendation == v) %>%
      arrange(publication_year)
    
    if(nrow(cat_v_data) < 5) next # Skip if not enough data points
    
    # Create Time Series
    data_ts <- ts(cat_v_data$percentage, start = min(cat_v_data$publication_year), frequency = 1)
    
    # Save Raw Data
    plot_data_raw <- rbind(plot_data_raw, data.frame(
      year = as.numeric(time(data_ts)),
      value = as.numeric(data_ts),
      variable = v,
      category = cat_name
    ))
    
    # --- Bai-Perron Modeling ---
    bp_model <- try(breakpoints(data_ts ~ 1, h = 0.15), silent = TRUE)
    if (inherits(bp_model, "try-error")) next
    
    opt_breaks <- if(!is.null(bp_model$BIC)) which.min(bp_model$BIC) - 1 else 0
    bp_final <- if(opt_breaks > 0) breakpoints(data_ts ~ 1, breaks = opt_breaks) else bp_model
    
    # Extract Fitted Trends
    plot_data_fitted <- rbind(plot_data_fitted, data.frame(
      year = as.numeric(time(data_ts)),
      value = as.numeric(fitted(bp_final)),
      variable = v,
      category = cat_name
    ))
    
    # Extract Breakpoints
    if (!is.null(bp_final$breakpoints) && any(!is.na(bp_final$breakpoints))) {
      plot_breaks <- rbind(plot_breaks, data.frame(
        break_year = as.numeric(time(data_ts)[bp_final$breakpoints]),
        variable = v,
        category = cat_name
      ))
    }
  }
}

# 6. Faceted Plot Generation
p <- ggplot() +
  geom_line(data = plot_data_raw, aes(x = year, y = value, color = variable), alpha = 0.3) +
  geom_line(data = plot_data_fitted, aes(x = year, y = value, color = variable), linewidth = 1.2) +
  geom_vline(data = plot_breaks, aes(xintercept = break_year, color = variable), 
             linetype = "dashed", alpha = 0.8) +
  facet_wrap(~category) +
  scale_color_manual(values = c("reject" = "#e74c3c", "approve" = "#2ecc71", "approve-with-reservations" = "#3498db")) +
  labs(title = "Reviewer Recommendation Trends by Research Category",
       y = "Percentage of Total Yearly Submissions (%)", x = "Year") +
  theme_minimal() + theme(legend.position = "bottom")

print(p)
```


## Causal Impact Analysis (stylistic constructs)
```{r}
# =========================================================
# F1000 AI-impact analyses (R)
# CausalImpact on AI-detection shares pre/post ChatGPT
# =========================================================

# ---- 1. Inputs ----
ts_path <- "/Users/Downloads/csv_files/f1000_scores.csv"

if (!file.exists(ts_path)) {
  stop("Missing f1000_scores.csv")
}

ts <- read_csv(ts_path, show_col_types = FALSE) %>%
  mutate(date = ymd(paste0(publication_year, "-07-01"))) %>%
  arrange(date)

# ---- 2. Intervention window ----
pre_period  <- c(min(ts$date), as.Date("2022-12-31"))
post_period <- c(as.Date("2023-01-01"), max(ts$date))

# =========================================================
# FDG causal
# =========================================================
if (!"share_flag_fdg" %in% names(ts)) stop("Missing share_flag_fdg")

z_fdg <- zoo(ts$share_flag_fdg, order.by = ts$date)
impact_fdg <- CausalImpact(z_fdg, pre.period = pre_period, post.period = post_period)

print(impact_fdg)
plot(impact_fdg)

png(
  filename = paste0(out_dir, "causal_impact_fdg.png"),
  width = 800,
  height = 600
)
plot(impact_fdg)
dev.off()


sink(paste0(out_dir, "causal_impact_summary_fdg.txt"))
cat(capture.output(summary(impact_fdg, "report")), sep = "\n")
sink()

# =========================================================
# LRR causal
# =========================================================
if (!"share_flag_lrr" %in% names(ts)) stop("Missing share_flag_lrr")

z_lrr <- zoo(ts$share_flag_lrr, order.by = ts$date)
impact_lrr <- CausalImpact(z_lrr, pre.period = pre_period, post.period = post_period)

print(impact_lrr)
plot(impact_lrr)

png(
  filename = paste0(out_dir, "causal_impact_lrr.png"),
  width = 800,
  height = 600
)
plot(impact_lrr)
dev.off()


sink(paste0(out_dir, "causal_impact_summary_lrr.txt"))
cat(capture.output(summary(impact_lrr, "report")), sep = "\n")
sink()

# =========================================================
# ImBD causal
# =========================================================
if (!"share_flag_imbd" %in% names(ts)) stop("Missing share_flag_imbd")

z_imbd <- zoo(ts$share_flag_imbd, order.by = ts$date)
impact_imbd <- CausalImpact(z_imbd, pre.period = pre_period, post.period = post_period)

print(impact_imbd)
plot(impact_imbd)

png(
  filename = paste0(out_dir, "causal_impact_imbd.png"),
  width = 800,
  height = 600
)
plot(impact_imbd)
dev.off()


sink(paste0(out_dir, "causal_impact_summary_imbd.txt"))
cat(capture.output(summary(impact_imbd, "report")), sep = "\n")
sink()

```
## Causal Impact Analysis (reviewer recommendations)
```{r}
# =========================================================
# F1000 Recommendation Trends → CausalImpact Analyses
# Approve / Approve-with-reservations / Reject
# =========================================================

# ---- 1. Load and construct input data ----
csv_path <- "/Users/Downloads/f1000_unique_scored.csv"

if (!file.exists(csv_path)) {
  stop("Input file not found: f1000_scored.csv")
}

# Read raw data
scored_all <- read_csv(csv_path, show_col_types = FALSE)

# Drop missing year or recommendation
scored_all <- scored_all %>%
  filter(!is.na(publication_year),
         !is.na(reviewer_recommendation))

# Count recommendations per year
trend <- scored_all %>%
  group_by(publication_year, reviewer_recommendation) %>%
  summarise(count = n(), .groups = "drop")

# Wide format: one column per recommendation type
trend_wide <- trend %>%
  pivot_wider(
    names_from  = reviewer_recommendation,
    values_from = count,
    values_fill = 0
  )

# Convert counts to yearly percentages
trend_percent <- trend_wide %>%
  mutate(total = rowSums(across(-publication_year))) %>%
  mutate(across(
    -c(publication_year, total),
    ~ .x / total * 100
  )) %>%
  select(-total)

# ---- 2. Prepare time series ----
ts_data <- trend_percent %>%
  mutate(date = ymd(paste0(publication_year, "-07-01"))) %>%
  arrange(date)

# ---- 3. Intervention window ----
pre_period  <- c(min(ts_data$date), as.Date("2022-12-31"))
post_period <- c(as.Date("2023-01-01"), max(ts_data$date))

# =========================================================
# Approve
# =========================================================
if (!"approve" %in% names(ts_data)) stop("Missing approve column")

z_approve <- zoo(ts_data$approve, order.by = ts_data$date)
impact_approve <- CausalImpact(z_approve, pre_period, post_period)

print(summary(impact_approve))

png(paste0(out_dir, "causal_impact_approve.png"), 800, 600)
plot(impact_approve)
dev.off()

sink(paste0(out_dir, "causal_impact_approve_summary.txt"))
cat(capture.output(summary(impact_approve, "report")), sep = "\n")
sink()

# =========================================================
# Approve with reservations
# =========================================================
if (!"approve-with-reservations" %in% names(ts_data)) {
  stop("Missing approve-with-reservations column")
}

z_res <- zoo(
  ts_data$`approve-with-reservations`,
  order.by = ts_data$date
)
impact_res <- CausalImpact(z_res, pre_period, post_period)

print(summary(impact_res))

png(
  paste0(out_dir, "causal_impact_approve_with_reservations.png"),
  800, 600
)
plot(impact_res)
dev.off()

sink(paste0(out_dir, "causal_impact_approve_with_reservations_summary.txt"))
cat(capture.output(summary(impact_res, "report")), sep = "\n")
sink()

# =========================================================
# Reject
# =========================================================
if (!"reject" %in% names(ts_data)) stop("Missing reject column")

z_reject <- zoo(ts_data$reject, order.by = ts_data$date)
impact_reject <- CausalImpact(z_reject, pre_period, post_period)

print(summary(impact_reject))

png(paste0(out_dir, "causal_impact_reject.png"), 800, 600)
plot(impact_reject)
dev.off()

sink(paste0(out_dir, "causal_impact_reject_summary.txt"))
cat(capture.output(summary(impact_reject, "report")), sep = "\n")
sink()

```

